## Hubble model {-}

Where N is the number of species in the local community, M the number of species in the metacommunity (set of communities), I is the number of individuals in the community, m is the immigration rate (the probability that a dead individuum will be replaced by an individuum from the metacommunity), d is the number of deaths at each time step, 

```{r}
pert<-perturbation(times = rep(500, N), durations = rep(100, N),
                   deathrate = N)
tsplot(simHubbell(N=N, M=N,I=1500,d=10, m=0.1, tskip=500, tend=1500,
                  perturb = pert))
```

Let's iterate the model n times

```{r}
Hubbell_tables<-list()
for(i in 1:50){
  Hubbell_tables[[i]]<-simHubbell(N=N, M=N,I=1500,d=10, m=0.1,
                                  tskip=500, tend=1500, perturb = pert)
}
tsplot(Hubbell_tables[[2]])
```

Diversity types
```{r}
hubbell_shannon<-ab_tables_div(Hubbell_tables, "shannon")
hubbell_pielou<-ab_tables_div(Hubbell_tables, "pielou")
hubbell_simpson<-ab_tables_div(Hubbell_tables, "simpson")
plot(hubbell_shannon[[6]], xlab = "Time",
     ylab = "Diversity", pch = 20)
plot(hubbell_pielou[[6]], xlab = "Time",
     ylab = "Evenness", pch = 20)
plot(hubbell_simpson[[6]], xlab = "Time",
     ylab = "Dominance", pch = 20)
```

Standard deviation of the diversity.
```{r}
plot(models_sd(Hubbell_tables), type = "l", lwd = 3, col = "brown",
     xlab = "Time steps", ylab = "sd")

models_sd<-function(ab_tables){
   div_list<-list()
 for(i in 1:length(ab_tables)){
   div_list[[i]]<-ab_table_div(ab_tables[[i]], "shannon")
 }
 div_mat<-matrix(unlist(div_list), ncol = 30)
 mat_sd<-c()
 for(i in 1:nrow(div_mat)){
   mat_sd[i]<-sd(div_mat[i,])
 }
 return(mat_sd)
}
```

