# Theorical ecology {-}

## Dynamics of microbial comunities {-}

Let's consider an hipothetical population that grows at an $ r $ rate: $ \dot{x} = rx $. If $ r>0 $, the population will increase exponentially $ x(t) = x_{0}e^{rt} $, a model that doens't make any biological sense.

```{r}
library(deSolve)
exponential <- function (time, y, parms) {
  with(as.list(c(y, parms)), {
  dN <- r * N
  list(dN)
 })
}
y <- c(N = 0.1)
parms <- c(r = 0.1, K = 1)
times <- seq(0, 100, 1)
plot(ode(y, times, exponential, parms))
```

To solve the previous problem, we add intraespecific competition for resources between the members of the population $ x^2 $ and a limit of aviable resources, represented as the medium carrying capasity.

```{r}
model <- function (time, y, parms) {
  with(as.list(c(y, parms)), {
  dN <- r * N * (1 - N / K)
  list(dN)
 })
}
y <- c(N = 0.1)
parms <- c(r = 1, K = 5)
times <- seq(0, 10, 1)
plot(ode(y, times, model, parms))
```

But the previous model is only considering an isolated population that has no interaction with any other species. The classical interaction model is the Lotka-Volterra. A generalization of this model is the so called generalized LV.

For simplify this problem, we'll consider only two species competing for resources (nutrients, space, water, etc.).

```{r}
# Lotka-Volterra model
LV2 <- function (time, y, parms) {
  with(as.list(c(y, parms)), {
  dN1 <- N1*(r1 - a11*N1 - a12*N2)
  dN2 <- N2*(r2 - a22*N2 - a21*N1)
  return(list(c(dN1, dN2)))
 })
}

# Parameters
coexitence <- c(r1 = 2, r2 = 2, a11 = 2.2 ,a22 = 2, a12 = 1.2, a21 = 1)
exclusionN1 <- c(r1 = 2, r2 = 2, a11 = 2.2 ,a22 = 1.2, a12 = 2, a21 = 1)
exclusionN2 <- c(r1 = 2, r2 = 2, a11 = 1.2 ,a22 = 2, a12 = 1, a21 = 2.2)
exclusionr1 <- c(r1 = 1, r2 = 2, a11 = 1 ,a22 = 1, a12 = 2, a21 = 2)
exclusionr2 <- c(r1 = 2, r2 = 1, a11 = 1 ,a22 = 1, a12 = 2, a21 = 2)

# Initial conditions and time steps
initial_conditions <- c(N1 = 2, N2 = 2)
time <- seq(0, 100, by = 1)

# Simulations
coex <- ode(initial_conditions, time, LV2, coexitence)
exN1 <- ode(initial_conditions, time, LV2, exclusionN1)
exN2 <- ode(initial_conditions, time, LV2, exclusionN2)
exr1 <- ode(initial_conditions, time, LV2, exclusionr1)
exr2 <- ode(initial_conditions, time, LV2, exclusionr2)

# Coexistence of both populations
matplot(x = coex[ , 1], y = coex[ , 2:3], type = "l",
        xlab = "Time", ylab = "Population",
        main = "Coexistence", lwd = 2,lty=1)
legend("topright", c("N1", "N2"), col = 1:2,lty=1)

# N2 excludes N1
matplot(x = exN1[ , 1], y = exN1[ , 2:3], type = "l",
        xlab = "Time", ylab = "Population",
        main = "N2 excludes N1", lwd = 2,lty=1)
legend("topright", c("N1", "N2"), col = 1:2,lty=1)

# N1 excludes N2
matplot(x = exN2[ , 1], y = exN2[ , 2:3], type = "l",
        xlab = "Time", ylab = "Population",
        main = "N1 excludes N2", lwd = 2,lty=1)
legend("topright", c("N1", "N2"), col = 1:2,lty=1)

# N2 excludes N1 by initial conditions
matplot(x = exr1[ , 1], y = exr1[ , 2:3], type = "l",
        xlab = "Time", ylab = "Population",
        main = "r1<r2", lwd = 2,lty=1)
legend("topright", c("N1", "N2"), col = 1:2,lty=1)

# N1 excludes N2 by initial conditions
matplot(x = exr2[ , 1], y = exr2[ , 2:3], type = "l",
        xlab = "Time", ylab = "Population",
        main = "r1>r2", lwd = 2,lty=1)
legend("topright", c("N1", "N2"), col = 1:2,lty=1)
```

## Generalized Lotka-Volterra model {-}

If we consider a community of N species, we use the generalized Lotka-Volterra model, that represents all possible interactions (positive, negative and neutral) between the species.

$$ \frac{dx_{i}}{dt} = x_{i} (r - \alpha_{ii}x_{i} - \sum_{j = 1; j \neq i}^{N}{\alpha_{ij}x_{j}}) $$

First, we must build an interaction matrix that represents the interactions between all species of the community.

```{r}
int_matrix<-function(N, aii, nep, pop){
   # Number of neutral interactions
   nep<-round(N*N*nep, 0)
   # Number of positive interactions
   pop<-round(N*N*pop, 0)
   # Stochastic negative interactions
   v<-runif(N*N, -1, 0)
   # Positive interactions
   positives<-v[sample(1:length(v), pop)]
   v[which(v %in% positives)]<-v[which(v %in% positives)]*(-1)
   # Negative interactions
   v[sample(1:length(v), nep)]<-0
   v[seq(1, length(v), N+1)]<-aii*(-1)
   # Interaction matrix
   mat<-matrix(v, N, N)
   return(mat)
}
# Let's apply the function
mat<-int_matrix(N = 10, aii = 0.5, nep = 0.4, pop = 0.2)
dim(mat)
```

Now we can simulate the system

```{r}
# Number of species
N<-ncol(mat)
# Growing rates
ri<-runif(N)
# Initial abundances
y<-runif(N)
# Parameters
parms<-cbind(ri, mat)
# Time steps
times<-seq(1, 100, by=1)
# gLV function
glv_solver<-function(t, y, parms){
  N<-nrow(parms)
  ri<-parms[,1]
  mat<-parms[,-1]
  glv<-y*(ri + mat %*% y)
  list(glv)
}
# Let's solve the function
sim_glv<-ode(y, times, glv_solver, parms)
# Time serie plot
sim_glv<-t(sim_glv[,-1])
colnames(sim_glv)<-times
matplot(x = times, y = t(sim_glv), type = "l",
        xlab = "Time", ylab = "Population",
        main = "Generalized Lotka-Volterra", lwd = 2,lty=1)
```

Let's add a perturbation to the system by changing the weights of interactions

```{r}
# We simulate another interaction matrix
mat_pert<-int_matrix(N = 10, aii = 0.5, nep = 0.4, pop = 0.2)
parms_pert<-cbind(ri, mat_pert)
# We simulate the model again using the new interaction matrix and the last abundances obtain in the previous simulation as initial abundances
times_pert<-seq(1, 30, by = 1)
sim_glv_pert<-ode(y = sim_glv[,ncol(sim_glv)], times=times_pert,
                    glv_solver, parms = parms_pert)
sim_glv_pert<-t(sim_glv_pert[,-1])
sim_glv_new<-cbind(sim_glv, sim_glv_pert)
# And now we re-establish the previous conditions
sim_glv_post<-ode(y = sim_glv_new[,length(times)], times=times,
                   glv_solver, parms = parms)
sim_glv_post<-t(sim_glv_post[,-1])
sim_glv_new<-cbind(sim_glv_new, sim_glv_post)

# Plot
matplot(x = seq(1:ncol(sim_glv_new)), y = t(sim_glv_new),
        type = "l", xlab = "Time", ylab = "Population",
        main = "Generalized Lotka-Volterra", lwd = 2,lty=1)
```

We compile the previous code into two functions

```{r}
# Generalized Lotka-Volterra function
my_glv<-function(mat, ri, y, t_start, t_end, perturb){
  require(deSolve)
  if(is.null(perturb)){
    # Number of species
    N<-ncol(mat)
    # We bind the growing rates and the interaction matrix
    parms<-cbind(ri, mat)
    # Time points
    times<-seq(t_start, t_end, by=1)
    # Simulation by using the glv_solver function
    sim_glv<-ode(y, times, glv_solver, parms)
    sim_glv<-t(sim_glv[,-1])
    colnames(sim_glv)<-times
    return(sim_glv)
  } else {
    N<-ncol(mat)
    parms<-cbind(ri, mat)
    # Community behavior before the perturbation
    times1<-seq(t_start, perturb[[2]]-1, by=1)
    sim_glv1<-ode(y, times1, glv_solver, parms)
    sim_glv1<-t(sim_glv1[,-1])
    # Community behavior during the perturbation
    times2<-seq(perturb[[2]], perturb[[3]], by=1)
    parms_p<-cbind(ri, perturb[[1]])
    # As initial abundances, we use the last abundances given by
    # the previous simulation 
    sim_glv2<-ode(sim_glv1[,ncol(sim_glv1)], times2, glv_solver, parms_p)
    sim_glv2<-t(sim_glv2[,-1])
    # We bind both simulations
    sim_glv2<-cbind(sim_glv1, sim_glv2)
    # Community behavior after the perturbation
    times3<-seq(perturb[[3]]+1, t_end, by=1)
    sim_glv3<-ode(sim_glv2[,ncol(sim_glv2)], times3, glv_solver, parms)
    sim_glv3<-t(sim_glv3[,-1])
    # We bind the final simulation and add the time points
    sim_glv_p<-cbind(sim_glv2, sim_glv3)
    final_time<-seq(1, ncol(sim_glv_p), by = 1)
    colnames(sim_glv_p)<-final_time
    return(sim_glv_p)
  }
}

# Perturbation function
perturb<-function(N, aii, nep, pop, strength, from, to){
   # We make another interaction matrix
   nep<-round(N*N*nep, 0)
   pop<-round(N*N*pop, 0)
   v<-runif(N*N, -1*(strength), 0)
   positives<-v[sample(1:length(v), pop)]
   v[which(v %in% positives)]<-v[which(v %in% positives)]*(-1)
   v[sample(1:length(v), nep)]<-0
   v[seq(1, length(v), N+1)]<-aii*(-1)
   mat_p<-matrix(v, N, N)
   # The function returns the matrix, and the starting and ending
   # perturbation time points
   return(list(mat_p, from, to))
}
```

Let's apply our functions

```{r}
N<-10
mat<-int_matrix(N, aii = 0.5, nep = 0.5, pop = 0.2)
prt<-perturb(N, aii = 0.5, nep = 0.5, pop = 0.2, strength = 0.5,
        from = 100, to = 130)
sim_1<-my_glv(mat, ri = runif(N), y = runif(N), t_start = 1, t_end = 250, NULL)
matplot(x = seq(1:ncol(sim_1)), y = t(sim_1),
        type = "l", xlab = "Time", ylab = "Population",
        main = "Generalized Lotka-Volterra", lwd = 2,lty=1)
sim_2<-my_glv(mat, ri = runif(N), y = runif(N), t_start = 1, t_end = 250, prt)
matplot(x = seq(1:ncol(sim_2)), y = t(sim_2),
        type = "l", xlab = "Time", ylab = "Population",
        main = "Generalized Lotka-Volterra", lwd = 2, lty=1)
```

## MacArthur model {-}

The MacArthur consumer-resource model analyses how the populations use the available resources, so Species interactions, therefore, aren't direct. Instead they emerge from the utilization of resources.

To simplify the model, we'll only consider two species competing for a single resource.

```{r}
dN1<-function(time, y, parms){
  dN1<-N1*Vl*C1l
}
```
